Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-06-26 paper/paper.Rmd
optimize intro

This cross-validation bias is well-known in spatial as well as non-spatial prediction [@brenning2005; @brenning2008; @arlot2010; @roberts2017] and in forecasting [@bergmeir2018].
It is most easily understood from a predictive modelling perspective by focusing on the question where (and when) the model should be used for prediction.
In crop classification from remotely-sensed data, for instance, learning samples routinely contain multiple grid cells from a sample of fields with known crop type, for instance 2000 grid cells from 100 fields scattered across a large study region.
The purpose of training a model on this sample is to make predictions on other, new *fields* within the same geographic domain (*intra-domain* prediction; @brenning2005) --- not *within* the same field, which obviously presents only one crop type.
In this situation it would therefore seem rather foolish to train a model on a simple random subsample of grid cells, and test it on the remaining data, i.e. using other grid cells from the same fields, as if we wanted to predict within a field.
The results from this performance assessment would be over-optimistic, and perhaps badly so.
To mimic the predictive situation for which the model is trained, we'd rather have to resample at the level of fields, not grid cells [@pena2015].
If the model was to be applied to adjacent agricultural districts, i.e. outside the learning sample's spatial domain (*extra-domain* prediction; @brenning2005), it would even seem necessary to resample at a higher level of spatial aggregation, i.e. at the level of agricultural districts within the learning sample, in order to realistically mimic the actual prediction task.
The cross-validation resampling needed therefore depends as much on the prediction task as on the data structure or dependency at hand.

@@

This toolbox, implemented as an open-source R package, builds upon and generalizes several existing toolboxes that have been developed in recent years for more specific settings (see Table \@ref(tab:sptcv-methods)).
The earliest and most comprehensive of these implementations is the \pkg{sperrorest} package [@brenning2012], which provides an extensible framework and includes predefined resampling strategies based on, for example, blocking, grouping and buffering.
In contrast, \pkg{blockCV} and \pkg{ENMeval} were developed for block and buffer resampling with a focus on species distribution modelling [@blockCV; @rest2014; @muscarella2014].
Neither of these have been integrated into established machine learning frameworks such as \pkg{mlr} or \pkg{caret}, and all of them lack support of temporal prediction tasks.
The \pkg{CAST} package, in contrast, focuses on spatiotemporal prediction tasks in a \pkg{caret} framework [@cast; @meyer2018].
One limitation of these packages is also that they focus on model assessment, while the proposed implementation within the \pkg{mlr} framework offers a seamless integration into model selection.
It is worth noting that a spatial cross-validation library \pkg{spacv} has also recently been developed for Python 3, which can be used with the \pkg{scikit-learn} machine learning framework [@pedregosa2011].

Thus, \pkg{mlr3spatiotempcv} implements for the first time a comprehensive state-of-the-art compilation of spatial and spatiotemporal partitioning schemes that is well-integrated into a comprehensive machine learning framework in R, the \pkg{mlr3} ecosystem [@mlr3].
This package is furthermore equipped with a variety of two- and three-dimensional diagram types.
Our hope is that this implementation will facilitate reproducible geospatial modelling and code-sharing across a broad range of application domains.

The purpose of this paper is to give an overview of the methods implemented in the R package \pkg{mlr3spatiotempcv}.
After presenting the conceptual background in the following section, we will outline the overall structure of the \pkg{mlr3spatiotempcv} package.
We will then compare and contrast the various spatial and spatiotemporal partitioning techniques, before we demonstrate their application in a machine learning model assessment in the following section.
We finally discuss implications and provide an outlook.

# Spatial and Spatiotemporal Cross-validation

>>>>

This cross-validation bias is well-known in spatial as well as non-spatial prediction [@brenning2005; @brenning2008; @arlot2010; @roberts2017] and in forecasting [@bergmeir2018].
It is most easily understood from a predictive modelling perspective by focusing on the question where (and when) the model should be used for prediction.
In crop classification from remotely-sensed data, for instance, learning samples routinely contain multiple grid cells from a sample of fields with known crop type, for instance 2000 grid cells from 100 fields scattered across a large study region.
The purpose of training a model on this sample is to make predictions on other, new fields within the same geographic domain (*intra-domain* prediction; @brenning2005) --- not *within* the same field, which obviously presents only one crop type.
In this situation it would therefore seem rather unwise to train a model on a simple random subsample of grid cells, and test it on the remaining data, i.e. using other grid cells from the same fields, as if we wanted to predict within a field.
The results from this performance assessment would be over-optimistic, and perhaps badly so.
To mimic the predictive situation for which the model is trained, one would rather have to resample at the level of fields, not grid cells [@pena2015].
If the model was to be applied to adjacent agricultural districts, i.e. outside the learning sample's spatial domain (*extra-domain* prediction; @brenning2005), it would even seem necessary to resample at a higher level of spatial aggregation, i.e. at the level of agricultural districts within the learning sample, in order to realistically mimic the actual prediction task.
The cross-validation resampling needed therefore depends as much on the prediction task as on the data structure or dependency at hand.

@@

This toolbox, implemented as an open-source R package, builds upon and generalizes several existing toolboxes that have been developed in recent years for more specific settings (see Table \@ref(tab:sptcv-methods)).
The earliest and most comprehensive of these implementations is the \pkg{sperrorest} R package [@brenning2012], which provides an extensible framework and includes predefined resampling strategies based on, for example, blocking, grouping and buffering.
In contrast, \pkg{blockCV} and \pkg{ENMeval} were developed for block and buffer resampling with a focus on species distribution modelling [@blockCV; @rest2014; @muscarella2014].
Neither of these however have been integrated into established machine learning frameworks such as \pkg{mlr}/\pkg{mlr3} or \pkg{caret}/\pkg{tidymodels}, and all of them lack support of temporal prediction tasks.
The \pkg{CAST} package, in contrast, focuses on spatiotemporal prediction tasks and makes use of some functions of the \pkg{caret} framework [@cast; @meyer2018].
One limitation of these packages is the sole focus on model assessment, while the proposed implementation within the \pkg{mlr3} framework offers a seamless integration into model selection.
It is worth noting that a spatial cross-validation library named \pkg{spacv} has also recently been developed for Python3, which can be used with the \pkg{scikit-learn} machine learning framework [@pedregosa2011].

Thus, \pkg{mlr3spatiotempcv} implements for the first time a comprehensive state-of-the-art compilation of spatial and spatiotemporal partitioning schemes that is well-integrated into a comprehensive machine learning framework in R, the \pkg{mlr3} ecosystem [@mlr3].
This package is furthermore equipped with a variety of two- and three-dimensional visualization capabilities.
Our hope is that this implementation will facilitate reproducible geospatial modeling and code-sharing across a broad range of application domains.

The purpose of this paper is to give an overview of the methods implemented in the R package \pkg{mlr3spatiotempcv}.
After presenting the conceptual background in the following section, the overall structure of the \pkg{mlr3spatiotempcv} package is outlined.
Next, various spatial and spatiotemporal partitioning techniques are contrasted and compared, before their application is demonstrated in a machine learning model assessment in the following section.
Finally, limitations of existing methods are discussed and an outlook is given.

# Spatial and spatiotemporal cross-validation
